\sysfull

### 2017-11-16 paper/evaluation.tex
evaluation section revised (compiler section)

We characterized the behavior and the efficiency of the \sys compiler. Table~\ref{table:compiler_result} shows, for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler
(column 10).

\begin{table}[t]
	\centering
	\footnotesize
        \renewcommand{\tabcolsep}{1pt}
	\begin{tabular}{|l|cc|cc|cc|cc|c|}
		\hline
		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
		\hline\hline
		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
		%dd &  &  &  &  &  &  &  & 287 &  \\
		\hline
	\end{tabular}
	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns). $^\dagger$dft does not compile with Clang}
\label{table:compiler_result}
\end{table}

On average, compiler-generated code introduces about 0.3\% more bytes of
protected data, suggesting that liveness analysis captures natural boundaries
in the use of variables also caught by a manual implementation.
Compiler-generated code generates about 16\% fewer tasks for applications, on
average.  The results indicate that the \sys compiler selects task boundaries
differently than a programmer would manually, favoring fewer, larger tasks.
The \sys compiler inserts protected data access operations conservatively when
a variable {\em may} alias a protected variables, due to the limitations of
pointer aliasing. Consequently, the \sys compiler introduces 48\% more
protected data accesses than programmer-written code. The run time results in
Section~\ref{sec:result_compiler_time} show that the increase in protected data
accesses sometimes degrades performance. Overall, the data show that the
compiler automatically generates tasks similarly to manually written tasks.
Doing so, \sys's compiler reaps the static task benefits of \sys, but
simplifies programming by sparing the programmer the job of task decomposition.

We measured compile time for \sys's compiler's automatic task generation for
each application\footnote{We could not compile the dft application since LLVM
inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3},
\texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not
provided by the LLVM MSP430 back-end.}. We compiled the applications on a
4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core
i5-6200U at 2.30\,GHz. We collected results using the Linux system time
collected in seconds. Most of the compile time was spent computing pointer
alias analysis, while calculating the possible live ranges of variables.
Applications such as RSA that have a larger code size and that manipulate more
variables have a longer compilation time, due to a larger search space for
pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter
than manually defining tasks, which requires hours of programmer time. The
characterization of \sys's compiler suggests that the \sys compiler enables
simpler task-based intermittent programming for \sys.

In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result},  
we experimentally assessed the performance of the \sys compiler, compared to the performance
of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}. 
%
Figure~\ref{fig:comp_user} compares the performance of manually written
benchmark code to the performance of code automatically compiled using \sys.
The figure shows run time for both cases, normalized to the run time of the
manually written program, running on continuous power.
%
The data show that in all cases except for ar and rsa, the run time performance
of compiled code generated by the \sys compiler has comparable performance to
manually written code.  With comparable performance in most cases, the data
suggest that the \sys compiler not only makes programming simpler, but does so,
in most cases, without a high run time overhead.  In the cases of both ar and
rsa, the code manipulates a large amount of data that is stored in an array and
passed into functions by reference.   We expect that the data being manipulated
by reference is leading to a large amount of may-alias references in the
compiler's analysis, which leads to a larger number of accesses to protected
data in the automatically compiled code, compared to the manually written code.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/comp_user}
	\caption{\sys compiler code overhead measured in terms of execution time per application (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
	\label{fig:comp_user}
\end{figure}

@@

>>>>

@@

\begin{table}[t]
	\centering
	\footnotesize
	\renewcommand{\tabcolsep}{1pt}
	\begin{tabular}{|l|cc|cc|cc|cc|c|}
		\hline
		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
		\hline\hline
		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
		%dd &  &  &  &  &  &  &  & 287 &  \\
		\hline
	\end{tabular}
	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang}\label{table:compiler_result}
\end{table}

\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}. 
%
%for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler (column 10).
%
On average, compiler-generated code introduces about 0.3\% more bytes of protected data, suggesting that liveness analysis captures natural boundaries in the use of variables also caught by a manual implementation. Compiler-generated code generates about 16\% fewer tasks for applications, on average. The results indicate that the \sys compiler selects task boundaries differently than a programmer would manually, favoring fewer, larger tasks. The \sys compiler inserts protected data access operations conservatively when a variable {\em may} alias a protected variables, due to the limitations of pointer aliasing. Consequently, the \sys compiler introduces 48\% more protected data accesses than programmer-written code. The run time results in Section~\ref{sec:result_compiler_time} show that the increase in protected data
accesses sometimes degrades performance. Overall, the data show that the compiler automatically generates tasks similarly to manually written tasks. Doing so, \sys's compiler reaps the static task benefits of \sys, but simplifies programming by sparing the programmer the job of task decomposition.

\textbf{Compile Time.} We measured compile time for \sys's compiler's automatic task generation for each application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}. We compiled the applications on a 4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core i5-6200U at 2.30\,GHz. We collected results using the Linux system time collected in seconds. Most of the compile time was spent computing pointer alias analysis, while calculating the possible live ranges of variables. Applications such as rsa that have a larger code size and manipulate more variables have a longer compilation time, due to a larger search space for pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter than manually defining tasks, which requires hours of programmer time. %The characterization of \sys's compiler suggests that the \sys compiler enables simpler task-based intermittent programming for \sys.

\textbf{Run Time of \sys Compiler Code.} In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result}, we experimentally assessed the performance of the \sys compiler, compared to the performance of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}. Figure~\ref{fig:comp_user} compares the performance of manually written benchmark code to the performance of code automatically compiled using \sys. The figure shows run time for both cases, normalized to the run time of the manually written program, running on continuous power. The data show that in all cases except for ar and rsa, the run time performance of compiled code generated by the \sys compiler has comparable performance to manually written code. With comparable performance in most cases, the data suggest that the \sys compiler not only makes programming simpler, but does so, in most cases, without a high run time overhead. In the cases of both ar and rsa, the code manipulates a large amount of data that is stored in an array and passed into functions by reference. We expect that the data being manipulated by reference is leading to a large amount of may-alias references in the compiler's analysis, which leads to a larger number of accesses to protected data in the automatically compiled code, compared to the manually written code.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/comp_user}
	\caption{\textbf{\sys compiler code overhead measured in terms of execution time per application} (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
	\label{fig:comp_user}
\end{figure}
